An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-30 draft.tex
Comments on comments
Comment out experiment 2

\subsection{Experiment 2}
\label{sec:exp2}
% Experiment 1

Having a multi-population based algorithm can decrease the execution time
because of the parallel execution of the evolution. But, having heterogeneous 
populations might enhance evolutionary search and needless
evaluations than homogeneous systems; heterogeneous settings, if
done right, increases the diversity of the whole population
\cite{araujo2008multikulti}.

But this is a rule of thumb, and it will depend on the degree of heterogeneity,
as well as on the algorithm itself. Some level of heterogeneity can be
implemented by just changing the configuration parameters of each population,
but in this case, we are interested in a heterogeneous search strategy. 
Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
BBOB testbed. We used ten populations and eight workers for the experiment and the
same parameters as before.

@@

We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. %What did you
                                %want to prove? - JJ
The left column has the
results when using five populations, and the right side shows the results for
ten. % Why only 5 and 10? - JJ
Charts show the execution time achieved by a certain number of workers.
Each box-plot indicates the time required for the execution of 15 function
instances, in seconds.  Each row has a different scale on the y-axis, to
compensate for the dimension increments. It is important to notice that the
execution time can also be affected by the algorithm finding a solution with an
error smaller than the threshold, before completing all the FEs.

As expected, in most cases, increasing the number of workers reduces the time
required for the execution of an instance. Mainly when using the same parameters
(number of populations and number of workers ). But, when we compare the two
options, we find some subtle differences: When using one or two
workers, using 5 populations is better in most cases. A 
reason for this could be that the migration of individuals takes more time to
reach the worker. For instance, when a single worker has 10 populations, only
after processing every one of them, it will start receiving mixed populations.  Although, in
all cases, and even a single worker, populations are heterogeneous, this
advantage takes more time to propagate. With four workers, almost the same
results can be achieved regardless of our selection between five or ten
populations. Also, there is not much difference between 4 and 8 workers, when
using five populations. In general, we can observe that keeping a ratio close to
1:1 between the number of workers and populations  better exploits the computing
resources available.
% I don't really understand this too well... JJ
% Also, it might just be that populations are waiting on the queue to
% be processed if there are more sub-populations than workers. The
% fact that 10 subpopulations take more with 1w than 5 subpopulations
% might be due to late startup as you say above, but you need to
% explain it better.
% Actually, what I see is no significant difference, or diminishing
% difference, between 5 and 10 sub-populations as the number of
% workers increases, and as long as the number of workers is less than
% the number of subpopulations.

\subsection{Experiment 2}
\label{sec:exp2results}

>>>>

% \subsection{Experiment 2}
% \label{sec:exp2}
% % Experiment 1

% Having a multi-population based algorithm can decrease the execution time
% because of the parallel execution of the evolution. But, having heterogeneous 
% populations might enhance evolutionary search and needless
% evaluations than homogeneous systems; heterogeneous settings, if
% done right, increases the diversity of the whole population
% \cite{araujo2008multikulti}.

% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
% as well as on the algorithm itself. Some level of heterogeneity can be
% implemented by just changing the configuration parameters of each population,
% but in this case, we are interested in a heterogeneous search strategy. 
% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
% BBOB testbed. We used ten populations and eight workers for the experiment and the
% same parameters as before.

@@

We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. %What did you
                                %want to prove? - JJ
The left column has the
results when using five populations, and the right side shows the results for
ten. % Why only 5 and 10? - JJ
     % mmm I really do not have an answer for that, I was thinking about 
     % testing from 1 to 15, and also more workers - Mario

Charts show the execution time achieved by a certain number of workers.
Each box-plot indicates the time required for the execution of 15 function
instances, in seconds.  Each row has a different scale on the y-axis, to
compensate for the dimension increments. It is important to notice that the
execution time can also be affected by the algorithm finding a solution with an
error smaller than the threshold, before completing all the FEs.

As expected, in most cases, increasing the number of workers reduces the time
required for the execution of an instance. Mainly when using the same parameters
(number of populations and number of workers ). But, when we compare the two
options, we find some subtle differences: When using one or two
workers, using 5 populations is better in most cases. A 
reason for this could be that the migration of individuals takes more time to
reach the worker. For instance, when a single worker has 10 populations, only
after processing every one of them, it will start receiving mixed populations.  Although, in
all cases, and even a single worker, populations are heterogeneous, this
advantage takes more time to propagate. With four workers, almost the same
results can be achieved regardless of our selection between five or ten
populations. Also, there is not much difference between 4 and 8 workers, when
using five populations. In general, we can observe that keeping a ratio close to
1:1 between the number of workers and populations  better exploits the computing
resources available.
% I don't really understand this too well... JJ
% Also, it might just be that populations are waiting on the queue to
% be processed if there are more sub-populations than workers. The
% fact that 10 subpopulations take more with 1w than 5 subpopulations
% might be due to late startup as you say above, but you need to
% explain it better.

% I was thinking also in terms of the heterogeneous populations.
% If population 1 to 6 are exploiting, and 7 to 9 are exploring
% they will take longer to catchup. 
% While with more workers this happens almost at the start
% If it makes sense, I will try to explain it better tomorrow with a 
% clear head - Mario

% Actually, what I see is no significant difference, or diminishing
% difference, between 5 and 10 sub-populations as the number of
% workers increases, and as long as the number of workers is less than
% the number of subpopulations.

% I think that was the point, but not too much less for instance 1w and 10pop
% Also I see a very big jump between 1w and 2w - Mario

% \subsection{Experiment 2}
% \label{sec:exp2results}

% \begin{figure*}[h!tb]
%     \begin{tabular}
%         {c@{\hspace*{-0.00001\textwidth}}
%          c@{\hspace*{-0.00001\textwidth}}
%          c@{\hspace*{-0.00001\textwidth}}
%         }
%     GA  &  PSO & GA \& PSO\\   
%     \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
%     \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
%     \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\

%     \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
%     \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
%     \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\

%     \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
%     \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
%     \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\

%     \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
%     \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
%     \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\

%     \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
%     \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
%     \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
%     \end{tabular}
%     \vspace{-3ex}
%      \caption{
%  Scaling of runtime with dimension to reach certain target values $\Delta f$. Lines:
%  average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
%  the most difficult target that was reached at least once (but not always);
%  Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
%  interquartile range with median of simulated runs; All values are divided by
%  dimension and plotted as $log_{10}$ values versus dimension. 
%  %Shown is the aRT for
%  %fixed values of $\Delta f = 10k$ with $k$ given in the legend. 
%  Numbers above aRT-symbols
%  (if appearing) indicate the number of trials reaching the respective target.
%  %The light thick line with diamonds indicates the best algorithm from BBOB 2009
%  %for the most difficult target. Horizontal lines mean linear scaling, slanted
%  %grid lines depict quadratic scaling.
%  }
%     \end{figure*}

%     \begin{figure}[h!tb]
%         \begin{tabular}
%             {c@{\hspace*{-0.00001\textwidth}}
%              c@{\hspace*{-0.00001\textwidth}}
%             }

%         \includegraphics[width=0.48\textwidth]{ppfigs_f001}&
%         \includegraphics[width=0.48\textwidth]{ppfigs_f002}\\

%         \includegraphics[width=0.48\textwidth]{ppfigs_f003}&
%         \includegraphics[width=0.48\textwidth]{ppfigs_f004}\\

%         \includegraphics[width=0.48\textwidth]{ppfigs_f005}&
%         \\
%         \end{tabular}
%         \vspace{-3ex}
%          \caption{Average running time (aRT in number of f-evaluations as $log_{10}$ value),
%           divided by dimension for target function value $10^{-8}$ versus dimension. 
%           Slanted grid lines indicate quadratic scaling with the dimension. 
%           Different symbols correspond to different algorithms given in the 
%           legend of $f_1$. Light symbols give the maximum number of 
%           function evaluations from the longest trial divided by dimension. 
%           Black stars indicate a statistically better result compared to 
%           all other algorithms with $p < 0.01$ and Bonferroni 
%           correction number of dimensions (six).}
%     \end{figure}
